---
title: '**Source code for data capture tool and data dictionary**'
author: "B210624"
date: "`r format (Sys.time (), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
[My assessment repository:](https://github.com/B210624/B210624_Working_with_data_types_and_structures_in_Python_and_R.git)# https://github.com/B210624/B210624_Working_with_data_types_and_structures_in_Python_and_R.git
## Load packages and data
Let’s load the packages and data needed for this script....
```{r a,message = FALSE, warning = FALSE}
library(tidyverse) #The *tidyverse* is a collection of R packages designed for  maniplating dataset..
library(here) #The *here* package is for easy filing in project-centered workflows.
library(knitr) #*knitr* is an R package that integrates code into text documents.
```
# **Overview**
In this assignment, I will use and load the Hospital length of stay (LOS) data (LOS_model) from the NHSRdatasets package. I will shortly see, investigate and tabulate the NHS Hospital length of stay (LOS) data set and save it to my RawData folder. I will search variable for later research as indicators for length of stay in hospital. Background: The data are csv.files from the NHSRdatasets package forskills development.
Hospital length of stay (LOS) data (LOS_model): Artificially generated hospital data. Fictional patients at ten fictional hospitals, with LOS, age and date status data. 

# **NHSRdatasets(Creation)**
```{r b,message = FALSE, warning = FALSE}
library(NHSRdatasets)
```
# **Store the NHS Hospital length of stay (LOS) data set (Storage)**
```{r c,message = FALSE, warning = FALSE}
data(LOS_model) #Load the LOS_model data.
write_csv(LOS_model, here("RawData", "los.csv")) 
#Here is the code to store theNHS Hospital length of stay (LOS) data set.
```
## Load the NHS Hospital length of stay (LOS) data set. Here is start of Synthesis
```{r d,message = FALSE, warning = FALSE}
LOS_CollectedData=read_csv(here("RawData", "los.csv"))
# I load the NHS Hospital length of stay (LOS) data (LOS_model) from RawData folder.
glimpse(LOS_CollectedData) #The `glimpse()` function is good to see the columns/variables in a data frame and show data type.
```
## Overview of my dataset of NHS Hospital length of stay (LOS) data (LOS_model)
I can see the los tibble consists of 300 rows of data and 5 columns with different classes. I have one order variable and four integer variables (or factors).
The dataset contains:     
* **ID:** All patients who stay in hospital has individual ID.  
*	**Organisation:** the Organisation is the fictional hospita; where the patients stay. 
*	**Age:** Age is patient age which is numeric data type and integer by using class 
* **Length of stay (LOS) :** the length of stay in hospital means how many days patients stay in hospital.    
*	**Death:** Death means the number of death in hospital.   

## Missing data
```{r e,message = FALSE, warning = FALSE}
#Calculate how many NAs there are in each variable.
LOS_CollectedData %>% map(is.na) %>%
map(sum) # 'map()' is a function for applying a function to each element of a list.
#The 'is.na' function produces a matrix, consisting of logical values.
```
The data is complete. I do not need to worry about manipulating missing data. 
### Build a data dictionary
for the data collected by the data capture tool. ##Build a linker data frame Firstly, I build a linker data frame. To do this, we need to create two string vectors representing the different variable descriptions and
types.
#Variable descriptions
```{r f,message = FALSE, warning = FALSE}
LOS_variable_description <- c("The ID column means that patient IDs which can be linked to the data collected to the original LOS_model data in the 'RawData' folder.",
"The organsation indicating that fictional hospital codes relates to the place where the patients stay.",
"The LOS showing length of staying in hosptail. It means how long days the patients stay in hospital.", 
"The death means that number of patient death as result of patients staying in hospital.",
"The Age meaning the patients age. It shows that age when the patients start staying in hosptail.")
print(LOS_variable_description)
```
###Variable types
I have three quantitative values (measured values) variables and one fixed values variables.
```{r g,message = FALSE, warning = FALSE}
glimpse(LOS_CollectedData) # I also used to see the data from LOS_CollectedData data frame.
```
I have three quantitative values (measured values) variables and one fixed values variables.
```{r h,message = FALSE, warning = FALSE}
LOS_variable_type <- c(0,1,0,0,0)
print(LOS_variable_type)
#I create a string vector representing the different variable types.
#It is a vector of integers with values 0 or 1. I need to use 0
#for a variable with quantitative values (measured values) variables and 1 for fixed values variables.
```
Now I use the build_linker() function from the dataMeta package to constructs an intermediary (linker) data frame between the CollectedData and the data dictionary. It requires the LOS_CollectedData data frame
and LOS_variable_description and LOS_variable_type string vectors.
```{r i,message = FALSE, warning = FALSE}
library(dataMeta) #Load the dataMeta to use build linker functon for the connection #for making dictionary.
linker<-build_linker(LOS_CollectedData,LOS_variable_description,LOS_variable_type)
print(linker)
```
###Making data dictionary
```{r j,message = FALSE, warning = FALSE}
LOS_dictionary <- build_dict(my.data = LOS_CollectedData, linker = linker)
#I use the build_dict() function from the dataMeta to constructs a data dictionary
#for a LOS_CollectedData data frame with the aid of the linker data frame between.
glimpse(LOS_dictionary) # Check the dictionary to see how looks like.
```

```{r k,message = FALSE, warning = FALSE}
# For my next task, it to save LOS_dictionary to my working data folder 'Data'
write_csv(LOS_dictionary, here("RawData", "LOS_CollectedData_DataDictionary.csv"))
```
### Append data dictionary to the CollectedData
As metadata, I now incorporate attributes to the LOS_CollectedData using the ‘incorporate_attr()’ function from the dataMeta package. The function requires the LOS_CollectedData and dictionary and main_string main_string as inputs.
```{r l,message = FALSE, warning = FALSE}
main_string <- "This data describes the NHS Hospital length of stay (LOS)
data set (LOS_model) from the *NHSRdatasets* package collected by the data capture tool."
main_string #Create main_string for attributes
```
## Incorporate attributes as metada
```{r m,message = FALSE, warning = FALSE}
#I use the ‘incorporate_attr()’ function to return an R dataset containing metadata stored in its attributes. The attributes include: a data dictionary, number of columns, number of rows, the name of the author and added it, the time when it was last edited, a brief description 
LOS_complete_CollectedData <- incorporate_attr(my.data = LOS_CollectedData, data.dictionary = LOS_dictionary,main_string = main_string)
#Change the author name
attributes(LOS_complete_CollectedData)$author[1]<-"B210624"
LOS_complete_CollectedData
```

```{r n,message = FALSE, warning = FALSE}
attributes(LOS_complete_CollectedData)
```

```{r o,message = FALSE, warning = FALSE}
save_it(LOS_complete_CollectedData, here("RawData", "LOS_complete_CollectedData"))
#I use the ‘save_it()’ function to save the LOS_CollectedData with attributes stored
# as metadata as an R dataset (.rds) into the ‘current working directory’RawData’ folder.
LOS_complete_CollectedData<-readRDS(here("RawData","LOS_complete_CollectedData.rds"))
# Here is the end of the process of Synthesis.
```