---
title: '**Descriptive Report**'
author: "B210624"
date: "`r format (Sys.time (), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
[My assessment repository:](https://github.com/B210624/B210624_Working_with_data_types_and_structures_in_Python_and_R.git)# https://github.com/B210624/B210624_Working_with_data_types_and_structures_in_Python_and_R.git
## Load packages and data
Let’s load the packages and data needed for this script....
```{r a,message = FALSE, warning = FALSE}
library(tidyverse) #The *tidyverse* is a collection of R packages designed for  maniplating dataset..
library(here) #The *here* package is for easy filing in project-centered workflows.
library(knitr) #*knitr* is an R package that integrates code into text documents.
library(scales)#*lubridate* provides tools that make it easier to manipulate dates in R. 
library(caret)# caret is a set of functions that attempt to streamline the process for creating predictive models.
```
# **Overview**
In this assignment, I will use and load the Hospital length of stay (LOS) data (LOS_model) from the NHSRdatasets package. I will shortly see, investigate and tabulate the NHS Hospital length of stay (LOS) data set and save it to my RawData folder. I will search variable for later research as indicators for length of stay in hospital. Background: The data are csv.files from the NHSRdatasets package forskills development.
Hospital length of stay (LOS) data (LOS_model): Artificially generated hospital data. Fictional patients at ten fictional hospitals, with LOS, age and date status data. 

# **NHSRdatasets(Creation)**
```{r b,message = FALSE, warning = FALSE}
library(NHSRdatasets)
```
# **Store the NHS Hospital length of stay (LOS) data set (Storage)**
```{r c,message = FALSE, warning = FALSE}
data(LOS_model) #Load the LOS_model data.
write_csv(LOS_model, here("RawData", "los.csv")) 
#Here is the code to store theNHS Hospital length of stay (LOS) data set.
```
## Load the NHS Hospital length of stay (LOS) data set. Here is start of Synthesis
```{r d,message = FALSE, warning = FALSE}
LOS_CollectedData=read_csv(here("RawData", "los.csv"))
# I load the NHS Hospital length of stay (LOS) data (LOS_model) from RawData folder.
glimpse(LOS_CollectedData) #The `glimpse()` function is good to see the columns/variables in a data frame and show data type.
class(LOS_CollectedData)#Look at the data class
head(LOS_CollectedData)#Look at data on the top 
tail(LOS_CollectedData)#Look at data on the bottom  
```
## Overview of my dataset of NHS Hospital length of stay (LOS) data (LOS_model)
I can see the los tibble consists of 300 rows of data and 5 columns with different classes. I have one order variable and four integer variables (or factors).
The dataset contains:     
* **ID:** All patients who stay in hospital has individual ID.  
*	**Organisation:** the Organisation is the fictional hospita; where the patients stay. 
*	**Age:** Age is patient age which is numeric data type and integer by using class 
* **Length of stay (LOS) :** the length of stay in hospital means how many days patients stay in hospital.    
*	**Death:** Death means the number of death in hospital.   

## Missing data
```{r e,message = FALSE, warning = FALSE}
#Calculate how many NAs there are in each variable.
LOS_CollectedData %>% map(is.na) %>%
map(sum) # 'map()' is a function for applying a function to each element of a list.
#The 'is.na' function produces a matrix, consisting of logical values.
```
The data is complete. I do not need to worry about manipulating missing data. 

## Let's tablulate the raw data for your report
```{r p,message = FALSE, warning = FALSE}
LOS_CollectedData %>%
  # Set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(LOS, Age, Death), comma) %>%
  # Show the first 10 rows
  head(10) %>%
  # Format as a table
  kable()
```

## Let's save the raw LOS_CollectedData to your 'RawData' folder
```{r q,message = FALSE, warning = FALSE}
write_csv(LOS_CollectedData, here("RawData", "LOS_CollectedData.csv"))
```

# **Selecting variables for your data capture tool**
```{r r,message = FALSE, warning = FALSE}
LOS_CollectedData_parameters<-LOS_CollectedData%>% select(ID, Organisation, Age, LOS)
```

## Let's tablulate the raw data for your report
```{r s,message = FALSE, warning = FALSE}
LOS_CollectedData_parameters%>%
  # Set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(LOS, Age), comma) %>%
  # Show the first 10 rows
  head(10) %>%
  # Format as a table
  kable()
```

### Let's visualise the relationship between LOS and Age before we save or raw data file.
```{r u,message = FALSE, warning = FALSE}
ggplot(LOS_CollectedData_parameters, aes(x=Age, y=LOS))  +
  facet_wrap(~Organisation, nrow=2)+
  geom_point() + geom_smooth() +
  labs(x = "Age",
       y = "LOS",
       title = "NHS Hospital length of stay (LOS) data set",
       caption = "Source: NHSRdatasets")
```

## Let's tablulate the raw data for your report
```{r v,message = FALSE, warning = FALSE}
LOS_CollectedData_parameters %>%
  # Set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(LOS, Age), comma) %>%
  # Show the first 10 rows
  head(10) %>%
  # Format as a table
  kable()
```

### Let's save provisional subsetted LOS_CollectedData to the 'RawData' folder
```{r w,message = FALSE, warning = FALSE}
glimpse(LOS_CollectedData_parameters)
write_csv(LOS_CollectedData_parameters, here("RawData", "LOS_CollectedData_parameters.csv"))
```

## Separating provisional ae_attendances_ENG_4hr_perfom data into training and testing sets
### How many rows are in the LOS_CollectedData?
```{r x,message = FALSE, warning = FALSE}
#The ae_attendances_ENG_4hr_perfom dataset is large with 
nrow(LOS_CollectedData_parameters) #rows of data
```

```{r y,message = FALSE, warning = FALSE}
LOS_prop<-(1-(15/nrow(LOS_CollectedData_parameters)))
#The proportion of the raw that needs to be assigned to the training data to ensure there is only 10 to 15 records in the test data is: 
print(LOS_prop)
```
##I use the createDataPartition() function from the *caret* package to splint our raw data into test and training data sets.
```{r z,message = FALSE, warning = FALSE}
set.seed(333)
#Partitioning the raw data into the test and training data.
LOS_trainIndex <- createDataPartition(LOS_CollectedData_parameters$ID, p = LOS_prop, 
                                  list = FALSE, 
                                  times = 1)
head(LOS_trainIndex)
# All records that are in the trainIndex are assigned to the training data.
LOS_Train <- LOS_CollectedData_parameters[ LOS_trainIndex,]
nrow(LOS_Train)
```
There are 288 records in my training data. 

### Let's tabulate LOS_CollectedData_parameters training data for your report
```{r aa,message = FALSE, warning = FALSE}
LOS_Train %>%
  # set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(LOS, Age), comma) %>%
  # show the first 10 rows
  head(10) %>%
  # format as a table
  kable()
  ```
  
### Our next task, it to save LOS_CollectedData_parameters training data to your working data folder 'Data'
```{r ab,message = FALSE, warning = FALSE}
write_csv(LOS_Train, here("Data", "LOS_CollectedData_parameters_train.csv"))
```

### Let's extract the LOS_CollectedData_parameters test data
#All records that are not in the LOS_trainIndex (`-trainIndex`) are assigned to the test data.
```{r ac,message = FALSE, warning = FALSE}
LOS_Test  <- LOS_CollectedData_parameters[-LOS_trainIndex,]
nrow(LOS_Test)
```
There are 12 records in my test data.
```{r ad,message = FALSE, warning = FALSE}
LOS_TestMarker  <- LOS_Test[1,]
```
#### Let's tabulate LOS_CollectedData_parameters marker test data for your report
```{r ae,message = FALSE, warning = FALSE}
LOS_TestMarker  %>%
  # set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(LOS, Age), comma) %>%
  # show the first 10 rows
  head(10) %>%
  # format as a table
  kable()
```
### Our next task, it to save my LOS_CollectedData_parametersmarker test data to our working data folder 'Data'
```{r af,message = FALSE, warning = FALSE}
write_csv(LOS_TestMarker, here("Data", "LOS_CollectedData_parameters_test_marker.csv"))
```

### We then need to set aside the remaining records for me to test (or collect with my) mydata-capture tool.
```{r ag,message = FALSE, warning = FALSE}
LOS_Test  <- LOS_Test[2:nrow(LOS_Test),]
```
#### Let's tabulate LOS_CollectedData_parameters test data for your report
```{r ah,message = FALSE, warning = FALSE}
LOS_Test  %>%
  # set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(LOS, Age), comma) %>%
  # show the first 10 rows
  head(10) %>%
  # format as a table
  kable()
```

### Our final task, is to save my LOS_CollectedData_parameters test data to our working data folder 'Data'
```{r ai,message = FALSE, warning = FALSE}
write_csv(LOS_Test, here("Data", "LOS_test.csv"))
```


